Understanding the Impact of Race, Region, and Employment on Voter Behavior: An Analysis of the 2020 U.S. Presidential Election

### 2024-03-16 paper/us_election_2020.qmd
final edits
some edits

# Introduction
Four years ago, the United States was at the forefront of an important electoral contest between the Democratic and Republican parties in the 2020 presidential election. The Democratic and Republican parties are represented by Donald Trump and Joe Biden respectively [@wiki2020]. Donald Trump achieved a remarkable victory in the 2016 election against Hillary Clinton [@result2016]. On the other hand, Joe Biden led the Democratic campaign. He aimed to regain control from the Republican party. In the end, Biden won the election by 51.3% majority votes [@cnn_2020], which Trump questioned its credibility. Due to the COVID-19 pandemic in 2020, the vote system changed to in-mail voting instead of in-person voting, which was why Trump believed that some votes were stolen [@trump_stolenvote]. Therefore, it is important for us to analyze why people voted for Biden and Trump in 2020 election and how various demographic, economic, and social factors may influence individuals' voting decisions. Perhaps then we could see why Biden eventually won. We analyze the 2020 election data set from Cooperative Election Study (CES) [@citeCes2020]. CES is a national online survey conducted before and after United States presidential elections [@wiki_ces]. We build a prediction model for individual voting preference using dataset from CES 2020 elections. This allows us to discover trends and address potential political, economic, and societal situations that may cause the eventual winning for Biden in 2020.

@@

Our discussion is made primarily based on the logistic regression model and its result in @tbl-modelresults1, as well as background information on different presidential candidates and their different views of the world and divergent approaches to tackling some of the  issues in the United States.

@@

## Credibility Interval {#sec-credibility-interval}

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: fig-modelresults1
#| fig-cap: "Credible intervals for predictors of support for Biden 1"

@@

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: fig-modelresults2
#| fig-cap: "Credible intervals for predictors of support for Biden 2"
# Create the model plot
model_plot <- modelplot(political_preferences1, conf_level = 0.9, , size = 0.2)

>>>>

# Introduction
Four years ago, the United States was at the forefront of an important electoral contest between the Democratic and Republican parties in the 2020 presidential election. The Democratic and Republican parties are represented by Donald Trump and Joe Biden respectively. Donald Trump achieved a remarkable victory in the 2016 election against Hillary Clinton. On the other hand, Joe Biden led the Democratic campaign. He aimed to regain control from the Republican party. In the end, Biden won the election by 51.3% majority votes [@cnn_2020], which Trump questioned its credibility. Due to the COVID-19 pandemic in 2020, the vote system changed to in-mail voting instead of in-person voting, which was why Trump believed that some votes were stolen [@trump_stolenvote]. Therefore, it is important for us to analyze why people voted for Biden and Trump in 2020 election and how various demographic, economic, and social factors may influence individuals' voting decisions. Perhaps then we could see why Biden eventually won. We analyze the 2020 election data set from Cooperative Election Study (CES) [@citeCes2020]. CES is a national online survey conducted before and after United States presidential elections [@wiki_ces]. We build a prediction model for individual voting preference using dataset from CES 2020 elections. This allows us to discover trends and address potential political, economic, and societal situations that may cause the eventual winning for Biden in 2020.

@@

Our discussion is made primarily based on the logistic regression model and its result in @tbl-modelresults1, as well as background information on the two presidential candidates, Trump and Biden, and their different views of the world and divergent approaches to tackling some of the issues in the United States.

@@

## Credibility Interval {#sec-credibility-interval}
@fig-modelresults1 and @fig-modelresults2 show the 90% credibility interval for the model.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: fig-modelresults1
#| fig-cap: "Credible intervals for predictors of support for Biden"

@@

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: fig-modelresults2
#| fig-cap: "Credible intervals for predictors of support for Biden with x_axis limits"
# Create the model plot
model_plot <- modelplot(political_preferences1, conf_level = 0.9, , size = 0.2)
